---
title: "Group Project"
author: 'SID: 450401242'
date: "18/10/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
abal <- read.csv(file="https://raw.github.sydney.edu.au/hlin4599/DATA2002-Assignment/master/abalone.csv?token=AAAG14_yU_uXZ2stJODFUYb_F56BCGUVks5b2Zi7wA%3D%3D")
library(GGally)
library(tidyverse)
library(dplyr)
require(plyr)
#install.packages("pairsD3")
#install_github("easyGgplot2", "kassambara")
#ggpairs(abal)
#glimpse(abal)
#shinypairs(abal)

#removing height outlier
colnames(abal)
abal = filter(abal, height != 0, height < 0.5)
abal = abal %>% mutate_at(.vars = vars(height), .funs = funs(ifelse(. == 0, NA, .))) %>% drop_na()
abal$rings = log(abal$rings)

#abalLog = select_if(abal, is.numeric)
#taking the log of the numeric data
#log(abalLog)
#adding sex category, might be out of order?
#abalLog$sex = abal$sex
#abal = abalLog

#summary(abal)



```

    
```{r}
require(ggplot2)
require(GGally)

#custom ggpairs graph
pm <- ggpairs(
 abal,
 mapping = ggplot2::aes(colour = sex),
 upper = list(continuous = wrap("density", alpha = 1), combo = "box"),
 lower = list(continuous = wrap("points" , size = 1, alpha = 1), 
              combo = wrap("dot", alpha = 0.4,size=0.2) ),
 title = "Rings"
)
#Making scatter plots, pulling them out of ggpairs

col = 9
pm[col,1]
pm[col,2]
pm[col,3]
pm[col,4]
pm[col,5]
pm[col,6]
pm[col,7]
pm


```

```{r}
library(tidyverse)
library(ggfortify)
library(GGally)
require(ggplot2)
require(caret)
abalone = abal
glimpse(abalone)

summary(abalone)

#simple linear model
model0_full = lm(rings ~ height, data = abalone)
summary(model0_full)

cv_model0 = train(
  rings ~ height, abal,
  method = "lm",
  trControl = trainControl(
    method = "cv", number = 10,
    verboseIter = FALSE
  )
)

cv_model0

#all predictors
cv_model1 = train(
  rings ~ ., abal,
  method = "lm",
  trControl = trainControl(
    method = "cv", number = 10,
    verboseIter = FALSE
  )
)
cv_model1

#plotting correlations of all variables

#multiple linear regression on all variables available
model1_full = lm(rings ~ ., data = abal)

#checking if any variables can be dropped using backwards stepwise method
data_back = step(model1_full, direction = "backward", trace = T)
summary(data_back)
autoplot(data_back, which = 1:2) + theme_bw()
data_back

cv_model1 = train(
  rings ~ sex + diameter + height + whole_weight + 
    shuckled_weight + viscera_weight + shell_weight, data = abal,
  method = "lm",
  trControl = trainControl(
    method = "cv", number = 10,
    verboseIter = FALSE
  )
)
cv_model1



#tried using the forwards and both direction stepwise method to drop variables, they didnt change the result significantly

#data_forward = step(model1_full, direction = "forward", trace = T)
#Multiple Linear Regression generated using forward step-wise method
#summary(data_forward)
#autoplot(data_forward, which = 1:2) + theme_bw()

#data_both = step(model1_full, direction = "both", trace = T)
#Multiple Linear Regression generated using step-wise method
#summary(data_both)
#autoplot(data_both, which = 1:2) + theme_bw()


#Multiple Linear Regression using an exhaustive search
library(leaps)
regsubset.out = regsubsets(rings ~ ., data = abalone, method = "exhaustive")
exhaustive_sum = summary(regsubset.out)
exhaustive = regsubset.out

plot(exhaustive_sum$rsq, xlab = "num variables", ylab = "R square", type = "b", main = "R Squared")
plot(exhaustive_sum$adjr2, xlab = "num variables", ylab = "adjusted r squared", type = "b", main = "Adjusted Rsquared ")
plot(exhaustive_sum$cp, xlab = "num variables", ylab = "complexity param", type = "b", main = "Complexity parameter")
plot(exhaustive_sum$bic, xlab = "num variables", ylab = "bic", type = "b", main = "BIC")


exhaustive_sum

library(caret)
cv_full = train(
  rings ~ sex + diameter + height + whole_weight + shuckled_weight + viscera_weight + shell_weight, abalone,
  method = "lm",
  trControl = trainControl(
    method = "cv", number = 10,
    verboseIter = FALSE
  )
)
cv_full

ex_model = (cv_full$finalModel)
autoplot(ex_model, which = 1:2)

colnames(abalone)
model2 = lm(rings ~ sex + length + diameter + height + shuckled_weight , data = abalone)
summary(model2)


#plotting the exhaustive model and the back stepwise model over each other, they're exactly the same... exhaustive model is red and dashed so you can see the blue line underneath



# suddenly not working?
#testPlot = pm[col,6]

#testPlot + geom_line(mapping = data_back)


colnames(abal)
#box = abal[,c("rings")] <- NULL
boxplot(abal, horizontal = TRUE, las = 1)

summary(ex_model)
cv_full

#confusionMatrix(ex_model, abal)


#Test performance other than cross validation???
#plot results???
#can we make the model better

# read dataset
df = mtcars


require(ggplot2)
require(reshape2)
abal2 = melt(abal, id.vars='rings')
ggplot(abal2) +
  geom_jitter(aes(value,rings, colour=variable)) + geom_smooth(aes(value,rings, colour="black"), method=lm, se=FALSE) +
  facet_wrap(~variable, scales="free_x") +
  labs(x = "Percentage cover (%)", y = "Number of individuals (N)")

ggplot(abal, aes(x = sex, y = rings)) + geom_point() + geom_smooth(method = lm, formula = data_back)

final <- lm(rings ~ diameter + height + whole_weight + shuckled_weight + viscera_weight + shell_weight, abal)

lm.mod<-function(df){
  p = lm(rings ~ sex + diameter, abal)
return(p) 
}

colnames(abal)

ggplot(abal, aes(x = diameter, y = rings)) + geom_point() + geom_smooth(, method = lm) 
ggplot(abal, aes(x = abal$whole_weight, y = rings)) + geom_point() + geom_smooth(, method = lm) 
ggplot(abal, aes(x = shuckled_weight, y = rings)) + geom_point() + geom_smooth(, method = lm)
ggplot(abal, aes(x = viscera_weight, y = rings)) + geom_point() + geom_smooth(, method = lm)

ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}

ggplotRegression(lm(rings ~ shuckled_weight + diameter + whole_weight + height  + viscera_weight + shell_weight, abal))


library(ggplot2)
theme_set(theme_classic())

m2 = lm(rings ~ whole_weight + length, data=abal)

pred.data = expand.grid(whole_weight = seq(min(abal$whole_weight), max(abal$whole_weight), length=5),length = quantile(abal$length))

pred.data$rings = predict(m2, newdata=pred.data)

ggplot(pred.data, aes(whole_weight, rings, colour=factor(length))) +
  geom_line() +
  labs(colour="height Quantiles") +
  labs(title = "rings ~ whole_weight + height")

library(caret)

control <- trainControl(method = "repeatedcv", 
                        number = 10, 
                        repeats = 6)

formula = rings ~ sex + diameter + height + whole_weight + 
    shuckled_weight + viscera_weight + shell_weight

model_lm <- train(formula, data = abal, method='lm',
                  trControl = control, tuneLength = 10)


#plot height
pred <- predict(model_lm, abal)
pred <- data.frame(pred = pred, height = abal$height)

library(reshape2)


heightplot <- ggplot(data = pred)+
  geom_point(aes(x = height, y = pred, colour = "Predicted Rings"), alpha = 0.3) +
  geom_point(data = abal, aes(x=height, y = rings, colour = "Observed Rings"), alpha = 0.3) +
  xlab("Height") +
  ylab("Number of Rings") +
  labs(title = "Comparing MLR by Height") +
  theme_dark()



#plot diameter

pred <- predict(model_lm, abal)
pred <- data.frame(pred = pred, diameter = abal$diameter)

diameterplot <- ggplot(data = pred)+
  geom_point(aes(x = diameter, y = pred, colour = "Predicted Rings"), alpha = 0.3) +
  geom_point(data = abal, aes(x=diameter, y = rings, colour = "Observed Rings"), alpha = 0.3) +
  xlab("Diameter") +
  ylab("Number of Rings") +
  labs(title = "Comparing MLR by diameter") +
  theme_dark()

#plot sex??

pred <- predict(model_lm, abal)
pred <- data.frame(pred = pred, sex = abal$sex)
library(cowplot)

sexplot1 <- ggplot(data = pred)+
  geom_boxplot(aes(x = sex, y = pred, colour = "Predicted Rings"), alpha = 0.3) +
  xlab("Sex") +
  ylab("Number of Rings") +
  labs(title = "Comparing MLR by sex") +
  theme_dark()

sexplot2 <- ggplot(data = pred)+
  geom_boxplot(data = abal, aes(x=sex, y = rings, colour = "Observed Rings"), alpha = 0.3) +
  xlab("Sex") +
  ylab("Number of Rings") +
  theme_dark()

sexplot3 <- plot_grid(sexplot1,sexplot2)

#plot whole weight

pred <- predict(model_lm, abal)
pred <- data.frame(pred = pred, whole_weight = abal$whole_weight)

wholeweightPlot <- ggplot(data = pred)+
  geom_point(aes(x = whole_weight, y = pred, colour = "Predicted Rings"), alpha = 0.3) +
  geom_point(data = abal, aes(x=whole_weight, y = rings, colour = "Observed Rings"), alpha = 0.3) +
  xlab("Whole Weight") +
  ylab("Number of Rings") +
  labs(title = "Comparing MLR by Whole Weight") +
  theme_dark()

wholeweightPlot

#plot shuck weight

pred <- predict(model_lm, abal)
pred <- data.frame(pred = pred, shuckled_weight = abal$shuckled_weight)

shuckWeightPlot <- ggplot(data = pred)+
  geom_point(aes(x = shuckled_weight, y = pred, colour = "Predicted Rings"), alpha = 0.3) +
  geom_point(data = abal, aes(x=shuckled_weight, y = rings, colour = "Observed Rings"), alpha = 0.3) +
  xlab("Shuckled Weight") +
  ylab("Number of Rings") +
  labs(title = "Comparing MLR by Shuckled Weight") +
  theme_dark()

shuckWeightPlot

#plot viscera weight

pred <- predict(model_lm, abal)
pred <- data.frame(pred = pred, viscera_weight = abal$viscera_weight)

visceraWeightPlot <- ggplot(data = pred)+
  geom_point(aes(x = viscera_weight, y = pred, colour = "Predicted Rings"), alpha = 0.3) +
  geom_point(data = abal, aes(x=viscera_weight, y = rings, colour = "Observed Rings"), alpha = 0.3) +
  xlab("Viscera Weight") +
  ylab("Number of Rings") +
  labs(title = "Comparing MLR by Viscera Weight") +
  theme_dark()



#plot shell weight

pred <- predict(model_lm, abal)
pred <- data.frame(pred = pred, shell_weight = abal$shell_weight)

shellWeightPlot <- ggplot(data = pred)+
  geom_point(aes(x = shell_weight, y = pred, colour = "Predicted Rings"), alpha = 0.3) +
  theme(legend.position="none") +
  geom_point(data = abal, aes(x=shell_weight, y = rings, colour = "Observed Rings"), alpha = 0.3) +
  theme(legend.position="none") +
  xlab("Shell Weight") +
  ylab("Number of Rings") +
  labs(title = "Comparing MLR by Shell Weight") +
  theme_dark()



shellWeightPlot
visceraWeightPlot
shuckWeightPlot
wholeweightPlot
diameterplot
heightplot
sexplot3

  


```

Constructing the model.
The data set had already been normalised by dividing all values by 200 but there were still some outliers and 0 values that we needed to remove. Obviously it doesn't make sense for an animal to have a height of 0 mm.

We then assessed the accuracy of a simple single linear regression model and the accuracy of a model that included all the terms.

We suspected we could find a better fit for the data so we used a back and forward stepwise method.

We also used an exhaustive search.

Notably all three methods returned the same model, which included all terms in the data set except for the length column.

-- Compare R squared, adjusted R squared and MAE

```{r}
#final full model
summary(ex_model)
cv_full

shellWeightPlot
visceraWeightPlot
shuckWeightPlot
wholeweightPlot
diameterplot
heightplot
sexplot3

#Multiple Linear Regression using an exhaustive search
library(leaps)
abalone2 = abalone %>% select(rings, height,whole_weight,diameter,length,sex)


regsubset.out2 = regsubsets(rings ~ ., data = abalone2, method = "exhaustive")
exhaustive_sum2 = summary(regsubset.out2)
exhaustive2 = regsubset.out

plot(exhaustive_sum2$rsq, xlab = "num variables", ylab = "R square", type = "b", main = "R Squared")
plot(exhaustive_sum2$adjr2, xlab = "num variables", ylab = "adjusted r squared", type = "b", main = "Adjusted Rsquared ")
plot(exhaustive_sum2$cp, xlab = "num variables", ylab = "complexity param", type = "b", main = "Complexity parameter")
plot(exhaustive_sum2$bic, xlab = "num variables", ylab = "bic", type = "b", main = "BIC")

exhaustive_sum2

library(caret)
cv_full = train(
  rings ~ sex + diameter + height + length + whole_weight, abalone2,
  method = "lm",
  trControl = trainControl(
    method = "cv", number = 10,
    verboseIter = FALSE
  )
)
summary(cv_full$finalModel)



```





